
---
title: ' Credit Scoring : Explanatory Data Analysis  '
author: "Yesmine Bellalah"
date: "Juin 2016"
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    fig_caption: yes
    fig_width: 8
    highlight: tango
    theme: cerulean
    toc: yes
---
#Credit Scoring : Exploratory Data Analysis

**Impoting training set**

```{r,echo=TRUE , warning=FALSE}

data=read.csv(file = file.choose(),row.names = 1)
View(data)

```

**Exploring data**

```{r, echo=TRUE, warning=FALSE}
summary(data)
str(data)

```


**Verification and visiualisation of  missing values**

```{r,echo=TRUE,warning=FALSE}
sapply(data,function(x) sum(is.na(x)) ) 
library(Rcpp)
library(Amelia)
missmap(data,main="missing values per variable",col = c("white","blue"))
```

**Converting ordinal variables to  factors to indicate that these variables should be treated as  categoricals variables** 

```{r,echo=TRUE, warning=FALSE}
for (i in c(2, 4:5, 7:13,15:21))
{data[,i]=as.factor(data[,i])}

```

**converting continuous variables to numeric**

```{r,echo=TRUE,warning=FALSE}
data$Age..years.=as.numeric(data$Age..years.)
data$Duration.of.Credit..month.=as.numeric(data$Duration.of.Credit..month.)
data$Credit.Amount=as.double(data$Credit.Amount)

```

**Separating factors and contiunous variables**

```{r,echo=TRUE,warning=FALSE}
quanti=data[,c(1,3,6,14)]
quali=data[,-c(1,3,6,14)]
```


#Univariate analysis

**1- Visualisation of the response variable Creditability**

```{r,echo=TRUE,warning=FALSE}
library(ggplot2)
library(stats)
library(base)
library(dplyr)
library(scales) 
d=as.data.frame(table(data$Creditability))
d
d$legend= c("Insolvent","Solvent")

blank_theme = 
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
        )
bp=ggplot(d, aes(x="", y=d$Freq, fill=d$legend))+
   geom_bar(width = 1, stat = "identity")

pie=bp + coord_polar("y", start=0)
target=pie + ggtitle("Creditworthiness") + blank_theme +
       theme(axis.text.x=element_blank())+
       geom_text(aes(y = d$Freq/2 + c(0, cumsum(d$Freq)[-length(d$Freq)]), 
       label = percent(d$Freq/sum(d$Freq))), size=5)+labs(fill="legend")
target

```

**2-Explanatory variables**

**2-1-continuous variables**

Density and histogram 

```{r,echo=TRUE,warning=FALSE}
library(ggplot2)
ggplot(data, aes(x=data$Age..years.)) + geom_histogram(aes(y=..density..), binwidth=.5, colour="black", fill="blue") + geom_density(alpha=.4, fill="#FF7667")  +geom_vline(aes(xintercept=mean(data$Age..years., na.rm=T)), color="red", linetype="dashed", size=1)

ggplot(data, aes(x=data$Duration.of.Credit..month.)) + geom_histogram(aes(y=..density..), binwidth=.5, colour="black", fill="blue") + geom_density(alpha=.4, fill="green")  +geom_vline(aes(xintercept=mean(data$Duration.of.Credit..month., na.rm=T)), color="red", linetype="dashed", size=1)

ggplot(data, aes(x=data$Credit.Amount)) + geom_histogram(aes(y=..density..), binwidth=.5, colour="black", fill="blue")  +geom_vline(aes(xintercept=mean(data$Credit.Amount, na.rm=T)), color="red", linetype="dashed", size=1)
ggplot(data, aes(x=data$Credit.Amount))+ geom_density(alpha=.4, fill="#FF7667")  +geom_vline(aes(xintercept=mean(data$Credit.Amount, na.rm=T)), color="red", linetype="dashed", size=1)
```

Dynamic Boxplots 
 
 
```{r,echo=TRUE,warning=FALSE}
library(plotly)
plot_ly(data,x=Age..years.,color=Creditability, type="box",boxpoints="all", jitter=0.4)
plot_ly(data,x=Duration.of.Credit..month.,color=Creditability, type="box",boxpoints="all", jitter=0.4)
plot_ly(data,x=Credit.Amount,color=Creditability, type="box",boxpoints="all", jitter=0.4)
```

**2-2-Factors**

Dynamic Pies 

```{r,echo=TRUE,warning=FALSE}
table(data$Account.Balance)
p1=plot_ly(data, labels = c(1,2,3), values = table(data$Account.Balance), type = "pie")
p2=layout(p1, title = "Account Balance ")
p2

table(data$Payment.Status.of.Previous.Credit)
p3=plot_ly(data, labels = c(1,2,3), values = table(data$Payment.Status.of.Previous.Credit), type = "pie")
p4=layout(p3, title = "Payment.Status.of.Previous.Credit")
p4

table(data$Purpose)
p5=plot_ly(data, labels = c(1,2,3,4), values = table(data$Purpose), type = "pie")
p6=layout(p5, title = "Purpose ")
p6

table(data$Value.Savings.Stocks)
p7=plot_ly(data, labels = c(1,2,3,4), values = table(data$Value.Savings.Stocks), type = "pie")
p8=layout(p7, title = "Value savings stocks ")
p8

table(data$Instalment.per.cent)
p9=plot_ly(data, labels = c(1,2,3,4), values = table(data$Instalment.per.cent), type = "pie")
p10=layout(p9, title = "Insatalment ")
p10

table(data$Sex...Marital.Status)
p11=plot_ly(data, labels = c(1,2,3), values = table(data$Sex...Marital.Status), type = "pie")
p12=layout(p11, title = "Sex")
p12

table(data$Guarantors)
p13=plot_ly(data, labels = c(1,2), values = table(data$Guarantors), type = "pie")
p14=layout(p13, title = "Guarontors ")
p14

table(data$Duration.in.Current.address)
p15=plot_ly(data, labels = c(1,2,3,4), values = table(data$Duration.in.Current.address), type = "pie")
p16=layout(p15, title = "Duration in current address")
p16

table(data$Most.valuable.available.asset)
p17=plot_ly(data, labels = c(1,2,3,4), values = table(data$Most.valuable.available.asset), type = "pie")
p18=layout(p17, title = "Most valuable available asset")
p18

table(data$Concurrent.Credits)
p19=plot_ly(data, labels = c(1,2), values = table(data$Concurent.Credits), type = "pie")
p20=layout(p19, title = "Concurrent.Credits")
p20

table(data$Type.of.apartment)
p21=plot_ly(data, labels = c(1,2,3), values = table(data$Type.of.apartment), type = "pie")
p22=layout(p21, title = "Type of apartment")
p22

table(data$No.of.Credits.at.this.Bank)
p23=plot_ly(data, labels = c(1,2), values = table(data$No.of.Credits.at.this.Bank), type = "pie")
p24=layout(p23, title = "No of credit in this bank ")
p24

table(data$Occupation)
p25=plot_ly(data, labels = c(1), values = table(data$Occupation), type = "pie")
p26=layout(p25, title = "Occupation ")
p26

table(data$No.of.dependents)
p27=plot_ly(data, labels = c(1,2), values = table(data$No.of.dependents), type = "pie")
p28=layout(p27, title = "No of dependents")
p28

table(data$Telephone)
p29=plot_ly(data, labels = c(1,2), values = table(data$Telephone), type = "pie")
p30=layout(p29, title = "Telephone")
p30

table(data$Foreign.Worker)
p31=plot_ly(data, labels = c(1,2), values = table(data$Foreign.Worker), type = "pie")
p32=layout(p31, title = "Foreign worker")
p32
```

#Visualisation of the response variable in function of the explanatory variables**

**Factors**

```{r,echo=TRUE,warning=FALSE}
plot(data$Account.Balance,data$Creditability)
plot(data$Purpose,data$Creditability)
plot(data$Payment.Status.of.Previous.Credit,data$Creditability)
plot(data$Value.Savings.Stocks,data$Creditability)
plot(data$Length.of.current.employment,data$Creditability)
plot(data$Instalment.per.cent,data$Creditability)
plot(data$Sex...Marital.Status,data$Creditability)
plot(data$Guarantors,data$Creditability)
plot(data$Duration.in.Current.address,data$Creditability)
plot(data$Most.valuable.available.asset,data$Creditability)
plot(data$Concurrent.Credits,data$Creditability)
plot(data$Type.of.apartment,data$Creditability)
plot(data$No.of.Credits.at.this.Bank,data$Creditability)
plot(data$No.of.dependents,data$Creditability)
plot(data$Telephone,data$Creditability)
plot(data$Foreign.Worke,data$Creditability)
```

**continuous variables** 

```{r,echo=TRUE,warning=FALSE}

ggplot(data, aes(x = data$Duration.of.Credit..month.)) +stat_density(aes(group = data$Creditability, color =data$Creditability ),position="identity",geom="line")
ggplot(data, aes(x = data$Age..years)) +stat_density(aes(group = data$Creditability, color =data$Creditability ),position="identity",geom="line")
ggplot(data, aes(x = data$Credit.Amount)) +stat_density(aes(group = data$Creditability, color =data$Creditability ),position="identity",geom="line")

```

#Correlation Matrix 

```{r,echo=TRUE,warning=FALSE}

library(corrplot)
N=cor(quanti)
corrplot(N,method = "number",type="upper")
library(corrgram)
corrplot.mixed(N, lower = "ellipse", upper = "number")

```




